RVD2: An ultra-sensitive variant detection model for low-depth targeted next-generation sequencing data

### 2013-11-27 doc/bioinformatics2013/RVD2.tex
Updated the manuscript

The HCC1187 cell line is derived from  epithelial cells from primary breast tissue from a 41 y/o adult with TNM stage IIA primary ductal carcinoma. A matched normal cells were derived from lymphoblastoid cells from whole blood. 
% TODO: check matched cells line origin
% double checked by Yuting, didn't find the matched cells line origin
% TODO: mention tumor purity here

Sequencing libraries were prepared from 500 ng of DNA using an early access version of Illuminas TruSeq DNA PCR-Free kit. The libraries were sequenced on four and eight lanes of a HiSeq 2000 flow cell for the normal and tumor, respectively, using 100 bp paired-end reads. Data, analyzed using the pre-release version of the Illumina Cancer Sequencing Workflow showed an average coverage of 40x for the normal and 90x for the tumor DNA with about 96\% coverage with 10 or more reads~\cite{}.
% TODO: write this in own words.
% TODO: add citation here.

% TODO: discuss alignment protocol

@@

@@

RVD2 identified five variants in the 44kbp PAXIP1 gene from chr7:XXX to chr7:XXX. Of the five called variants, three were homozygous germline variants and two were heterozygous germline variants. The two heterozygous variants had a loss-of-heterozygosity in the tumor sample.

Figure~\ref{fig:brca_MAF} shows the estimated minor allele frequencies for the normal and tumor samples at the called locations. Positions chr7:154743899C$>$T, chr7:154753635T$>$C and chr7:154780960C$>$T are germline homozygous mutations. Positions chr7:154754371 and chr7:154758813 are called heterozygous in the normal sample. In the tumor cells we identified all five mutations chr7:154743899C$>$T, chr7:154753635T$>$C, chr7:154754371T$>$C, chr7:154758813G$>$A, and chr7:154780960C$>$T. Positions chr7:154754371T$>$C and chr7:154758813G$>$A then loss-of-heterozygosity events.

%According to dbSNPv138 chr7:154743899 is rs1239326, chr7:154753635 is rs1239324, chr7:154754371 is rs71534174, chr7:154758813 is rs35505514, and chr7:154780960 is rs4398858.
%TODO: place these rs labels beneath the position labels on the x-axis of the figure. Remove the above sentence and add. 
%Some of these mutations are also found to be common population SNPs according to dbSNPv138. The corresponding rs identities are shown in the figure.

@@

>>>>

The HCC1187 cell line is derived from  epithelial cells from primary breast tissue from a 41 y/o adult with TNM stage IIA primary ductal carcinoma. A matched normal cells were derived from lymphoblastoid cells from whole blood. Sequecing libraries were prepared according to the protocal described in the original techical report~\cite{}. The raw FASTQ read files were aligned to hg19 igenome ~\cite{} using the Isaac aligner~\cite{} to generate BAM files. We used samtools mpileup to generate pileup files using hg19 igenome ~\cite{} as reference sequence.

% TODO:cite the paper Raczy C, Petrovski R, Saunders C, Chorny I, Kruglyak S, Margulies E, Chuang HY, Kallberg M, Kumar SA, Liao A, Little KM, Stromberg M, Tanner S. Isaac: Ultra-fast whole genome secondary analysis on Illumina sequencing platforms. 2013
% TODO:cite the technical report.
% TODO: check matched cells line origin
% double checked by Yuting, didn't find the matched cells line origin
% TODO: mention tumor purity here

%Sequencing libraries were prepared from 500 ng of DNA using an early access version of Illuminas TruSeq DNA PCR-Free kit. The libraries were sequenced on four and eight lanes of a HiSeq 2000 flow cell for the normal and tumor, respectively, using 100 bp paired-end reads. Data, analyzed using the pre-release version of the Illumina Cancer Sequencing Workflow showed an average coverage of 40x for the normal and 90x for the tumor DNA with about 96\% coverage with 10 or more reads~\cite{}.
% TODO: write this in own words.
% TODO: add citation here.

% TODO: discuss alignment protocol
% TODO: rvd2 parameters
% TODO: chi2 test
% TODO: Pubmed for peer-reviewed paper on HCC1187 sample

@@

Considering HCC1187 sample is diploid, we did two tests based on posterior density using RVD2. First, we compared the posterior density $\mu_j^{\Delta}$ to find out if $\mu_j$ in case sample was significantly higher than that in control sample and vice verse.Threshold $\tau$ was set at $1\times10^{-3}$ to avoid error brought by computer limitation and confidence level $1-\alpha$ was set at 0.95. % The difference is constant at 1.74*10^{-4}
\begin{equation}\label{eqn:bayes_test}
	\Pr( \tilde{\mu}_j^{\text{case}} - \tilde{\mu}_j^{\text{control}} \geq \tau ) > 1-\alpha,
\end{equation}
\begin{equation}\label{eqn:bayes_test}
	\Pr( \tilde{\mu}_j^{\text{control}} - \tilde{\mu}_j^{\text{case}} \geq \tau ) > 1-\alpha,
\end{equation}

%TODO: write up the somatic test section
%TODO: specify the results from somatic test or from the $\mu_j^{\Delta}$ posterior density test.

Second, we identified the gemerline status and somatic status of HCC1187 sample by looking into posterior density of control sample and case sample independently. We can compute
$[0.0,0.2],[0.2,0.8],[0.8.1]$
confidence level $1-\alpha$ was set at 0.8

\begin{equation}\label{eqn:bayes_test}
	\Pr( \tilde{\mu}_j \in [\tau_\text{l},\tau_\text{u}] ) > 1-\alpha,
\end{equation}

@@

RVD2 identified five variants in the 44kbp PAXIP1 gene from chr7:154738059 to chr7:154782774. Of the five called variants, three were homozygous germline variants and two were heterozygous germline variants. The two heterozygous variants had a loss-of-heterozygosity in the tumor sample.

Figure~\ref{fig:brca_MAF} shows the estimated minor allele frequencies for the normal and tumor samples at the called locations. Positions chr7:154743899C$>$T, chr7:154753635T$>$C and chr7:154780960C$>$T are germline homozygous mutations. Positions chr7:154754371 and chr7:154758813 are called heterozygous in the normal sample. In the tumor cells we identified all five mutations chr7:154743899C$>$T, chr7:154753635T$>$C, chr7:154754371T$>$C, chr7:154758813G$>$A, and chr7:154780960C$>$T. Positions chr7:154754371T$>$C and chr7:154758813G$>$A then loss-of-heterozygosity events.Some of these mutations are also found to be common population SNPs according to dbSNPv138. The corresponding rs identities are shown in the figure.

@@

% we ran strelka and it found three positions: chr7:154754371, chr7:154758813 and chr7:154760439.
